---
title: "Predicting respondent difficulty in web surveys: A machine-learning approach based on mouse movement features"
output: html_document
---

#### R codes method of correction "baseline and position"
#### Authors: Amanda Fernández-Fontelo, Pascal J. Kieslich, Felix Henninger, Frauke Kreuter, Sonja Greven


```{r libraries,echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(mousetrap)
library(ggplot2)
library(tidyr)
```

```{r data, echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
data=read.csv("...",header=TRUE,sep=";")
```

#### Position residuals

The code below fits the equation: 
$$Y_{ijk}=\gamma_{jk0}+\sum_{m=1}^{m_{k-1}}\gamma_{jkm}P_{ik}(m)+w_{ijk},$$
which corresponds to equation (3) in the manuscript.

```{r position, echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
# Number of baseline questions: 8
# Number of target questions  : 3
questions=unique(data$question)
questions=questions[questions!="employment_status"]
questions=questions[questions!="employment_changes_slider"]
mt_indices =  c(paste("xpos_flips_",c(0,4,10,20,50,100,200),sep=""),
                paste("ypos_flips_",c(0,4,10,20,50,100,200),sep=""),
                paste("hovers_",c(250,500,1000,2000,3000),sep=""),
                paste("hover_time_",c(250,500,1000,2000,3000),sep=""),
                "RT",
                "initiation_time",
                "idle_time",
                "total_dist",
                "vel_max",
                "vel_max_time",
                "acc_max",
                "acc_max_time",
                "acc_min",
                "acc_min_time")
data_residuals=data.frame()
for (question in questions){
        current_data=data %>% filter(question==!!question)
        for (var in mt_indices){
                current_lm=lm(as.formula(paste(var,"position",sep="~")),data=current_data)
                current_data=cbind(current_data,resid(current_lm))
                colnames(current_data)[ncol(current_data)]=paste(var,c("resid"),sep="_")}
        data_residuals=bind_rows(data_residuals,current_data)}
pos_corr_indices=data_residuals
```

#### Specify questions

```{r questions, echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
questions=unique(pos_corr_indices$question)
questions=questions[!questions%in%c("employee_level","demography_edu")]
# target question
questions_crit="employment_detail"
# baseline questions
questions_pred=questions[questions!=questions_crit]
```

#### Remove participants with missing data 

```{r missing, echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
# Selecting just individuals with responses in the 8 baseline variables + 3 target variables
id_obs=pos_corr_indices %>% 
       select(id,question,RT) %>% 
       filter(question %in% !!questions) %>% 
       mutate(RT=1) %>%
       spread(key="question",value="RT") %>% 
       filter(!is.na(questions_crit))
id_obs$n_data=rowSums(id_obs[,-1],na.rm=TRUE)
id_obs=filter(id_obs,n_data==9)
pos_corr_indices=filter(pos_corr_indices,id %in% id_obs$id)

# Computing the average within the same individual in the same index across the eight baseline questions
pos_corr_indices_crit=pos_corr_indices[!pos_corr_indices[,"question"]%in% 
                      c("employment_detail","demography_edu","employee_level"),]
idunique=unique(pos_corr_indices[,"id"])
results_pos_average=data.frame(id=idunique[1],
t(apply(pos_corr_indices_crit[pos_corr_indices_crit[,"id"]==idunique[1],49:82],2,mean)))
for(j in 2:length(idunique)){
        results_pos_average=rbind(results_pos_average,data.frame(id=idunique[j],
        t(apply(pos_corr_indices_crit[pos_corr_indices_crit[,"id"]==idunique[j],49:82],2,mean))))}
results=pos_corr_indices[pos_corr_indices[,"question"]=="employment_detail",c(1,49:82)]
names(results)[2:ncol(results)]=paste(names(results)[2:ncol(results)],"_empdetail",sep="")
results_pos_average=merge(results_pos_average,results,by="id")
```

#### Baseline residuals

The code below fits the equation: 
$$\widehat{w}_{ijk}=\phi_{j0}+\phi_{j1}\bar{\widehat{\omega}_{ij}}+v_{ijk},$$
which corresponds to equation (4) in the manuscript.

```{r final, echo=TRUE,eval=FALSE,warning=FALSE,message=FALSE}
mt_indices =  c(paste("xpos_flips_",c(0,4,10,20,50,100,200),sep=""),
                paste("ypos_flips_",c(0,4,10,20,50,100,200),sep=""),
                paste("hovers_",c(250,500,1000,2000,3000),sep=""),
                paste("hover_time_",c(250,500,1000,2000,3000),sep=""),
                "RT",
                "initiation_time",
                "idle_time",
                "total_dist",
                "vel_max",
                "vel_max_time",
                "acc_max",
                "acc_max_time",
                "acc_min",
                "acc_min_time")

corrected_indices=data.frame(x=residuals(lm(xpos_flips_0_resid_edulevel~xpos_flips_0_resid,
                                            data=results_pos_average)))
for(i in 2:length(mt_indices)){
        corrected_indices=cbind(corrected_indices,
        x=residuals(lm(results_pos_average[,i+35]~results_pos_average[,i+1])))}
corrected_indices=cbind(id=results_pos_average[,"id"],corrected_indices)
names(corrected_indices)[2:ncol(corrected_indices)]=paste(mt_indices,"_corrected",sep="")
data_residuals_baseline=merge(pos_corr_indices[pos_corr_indices[,"question"]=="employment_detail",1:48],
                              corrected_indices,by="id")
```

